Future challenges for Research Software Engineering

### 2025-05-12 future_challenges_for_RSEng.tex
finalise whatis boxes

\subsection{Definition and Evolution of RSEs}
\begin{whatis}{What do we want to have achieved}{ethics}
The term RSE is established in the minds of university boards, as a valuable member in the academic framework.
RSEs are developed using specialized master's program, and their support is valued by researchers.
\end{whatis}
% Here we consider how the topics of RSEng will change in the face of the digitalization of society.
% Also here: What will our topics be after version control doesn't need to be taught anymore?
% Digitalisation will also feature more heavily in the domain curricula.
In this discussions the workshop participants found that
the digital skills of domain scientists will increase,
and software will become easier to interact with,
reducing reliance on RSEs to train newcomers on the most basic tasks
(JN: for example, versioning and archiving?)
(GC: Higher-level and more intuitive workflows that require less training).
User training sessions might focus on more advanced topics
that cater to a smaller audience, or involve mentoring more frequently.

@@

\subsection{Development of the RSE community}
\begin{whatis}{What do we want to have achieved}{ethics}
We have more national RSE communities that are interacting with other partners in their nation.
There are regular international conferences.
\end{whatis}
In recent years, we saw the formation of RSE institutes and advocacy groups, such as
the UK Software Sustainability Institute\footnote{\url{https://www.software.ac.uk}},
the US Research Software Sustainability Institute\footnote{\url{https://urssi.us}},
the European Virtual Institute for Research Software Excellence\footnote{\url{https://everse.software}},
and the EOSC Opportunity Area Expert Group 7 on Research Software%
\footnote{\url{https://eosc.eu/opportunity-area-exp/oa7-research-software}}.
We expect these initiatives to help increase the visibility and accelerate
the professionalisation%
\footnote{MM: do we want to decide for UK or US English? Or keep it mixed? Not being a native English speaker I don't know how it reads.}
of the RSE role.
In addition, a multitude of RSE communities self-organised at the national
and regional level to support more viable RSE career paths in academia.
We can expect RSEs to become more easily accessible,
possibly through centrally-funded long-term positions at research institutions.
The community of RSE's will get a significant boost in their acceptance if RSE Masters programmes
become available at various universities.

@@

\begin{whatis}{What do we want to have achieved}{academic}
School curricula have embraced the digital transformation, and have included basics of computer interaction
in every schoolyear.
In Universities, the first generations of RSE-Master graduates find their ay into research departments,
and accelerate the pace of science.
\end{whatis}

>>>>

\subsection{Definition and Evolution of RSEs}
\begin{whatis}{What do we want to have achieved}{ethics}
The term RSE is established in the minds of university boards, as a valuable member in the academic framework.
RSEs are developed using specialized master's program, and their support is valued by researchers. We still teach researchers basic software skills, but the content of this lecture has changed drastically.
\end{whatis}
% Here we consider how the topics of RSEng will change in the face of the digitalization of society.
% Also here: What will our topics be after version control doesn't need to be taught anymore?
% Digitalisation will also feature more heavily in the domain curricula.
In this discussions the workshop participants found that
the digital skills of domain scientists will increase,
and software will become easier to interact with,
reducing reliance on RSEs to train newcomers on the most basic tasks
(JN: for example, versioning and archiving?)
(GC: Higher-level and more intuitive workflows that require less training).
User training sessions might focus on more advanced topics
that cater to a smaller audience, or involve mentoring more frequently.

@@

\subsection{Development of the RSE community}
\begin{whatis}{What do we want to have achieved}{ethics}
We have more national RSE communities that are interacting with other partners in their nation.
There are regular international conferences. The installation of RSE Master's programmes led to
increased numbers of the members of the national societies.
\end{whatis}
In recent years, we saw the formation of RSE institutes and advocacy groups, such as
the UK Software Sustainability Institute\footnote{\url{https://www.software.ac.uk}},
the US Research Software Sustainability Institute\footnote{\url{https://urssi.us}},
the European Virtual Institute for Research Software Excellence\footnote{\url{https://everse.software}},
and the EOSC Opportunity Area Expert Group 7 on Research Software%
\footnote{\url{https://eosc.eu/opportunity-area-exp/oa7-research-software}}.
We expect these initiatives to help increase the visibility and accelerate
the professionalisation%
\footnote{MM: do we want to decide for UK or US English? Or keep it mixed? Not being a native English speaker I don't know how it reads.}
of the RSE role.
In addition, a multitude of RSE communities self-organised at the national
and regional level to support more viable RSE career paths in academia.
We can expect RSEs to become more easily accessible,
possibly through centrally-funded long-term positions at research institutions.
The community of RSE's will get a significant boost in their acceptance if RSE Masters programmes
become available at various universities.

@@

\begin{whatis}{What do we want to have achieved}{academic}
School curricula have embraced the digital transformation, and have included basics of computer interaction
in every class level.
In universities, the first generations of RSE-Master graduates find their way into research departments,
and accelerate the pace of science.
\end{whatis}
